---
title: "PFIM: Paleo Food-web Inference Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PFIM: Paleo Food-web Inference Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# PFIM: Paleo Food-web Inference Model

`PFIM` is an R package for **inferring predator–prey interactions** using trait-based 
rules and generating **hypothetical realized food webs**. It implements 
the **Paleo Food-web Inference Model (Shaw 2024)** and includes tools for categorical 
trait matching, numerical predator–prey size rules, and power-law downsampling of 
realised webs.

---

## Installation

```{r eval=FALSE}
# Install from CRAN (when available)
install.packages("pfim.R")

# Or install development version from GitHub
# install.packages("devtools")
devtools::install_github("BecksLab/pfim.R")
```

## Load Example Data

```{r eval=FALSE}
library(pfim.R)

# Trait data
data("traits", package = "pfim.R")

# Trait combination rules
data("feeding_rules", package = "pfim.R")

head(traits)
head(feeding_rules)
```

## Inferring a consumer–resource edgelist

Use `infer_edgelist()` to infer feasible interactions:

```{r eval=FALSE}

edgelist <- infer_edgelist(
  data = traits,
  cat_combo_list = feeding_rules,
  col_taxon = "species",
  certainty_req = "all",
  allow_self = FALSE
)

# Show first few interactions
head(edgelist)

```

## Generate hypothetical realised webs

Create a series of hypothetical webs using `powerlaw_prey()`

```{r eval=FALSE}

webs <- powerlaw_prey(
  el = edgelist,
  n_samp = 5,
  y = 2.5
)

# Inspect first web
webs[[1]]

```

Each element of `webs` is a matrix representing a realised food web generated 
from the inferred interactions