---
title: "PFIM: Paleo Food-Web Inference Model"
author: "Your Name"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PFIM Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette demonstrates how to use the PFIM package to infer
consumer–resource interactions from trait data, and generate
hypothetical realized food webs using a power-law link distribution.

This workflow demonstrates:

1. Inferring feasible interactions with `infer_edgelist()`.
2. Downsampling interactions into hypothetical realized webs with `powerlaw_prey()`.
3. Customizing link distributions via the `func` argument.

Users can now use these realised webs for **network analyses**,  
simulation studies, or **comparisons to fossil and modern food webs**.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(pfim)

# Load example data included in the package
data("traits", package = "pfim")
data("feeding_rules", package = "pfim")
data("interactions", package = "pfim")  # internal edgelist for tests
```

## Infer Edgelist

Infer interactions using categorical trait rules

```{r infer-edgelist}

edgelist <- infer_edgelist(
  data = traits,
  cat_combo_list = feeding_rules,
  col_taxon = "species",
  certainty_req = "all",
  hide_printout = TRUE
)

head(edgelist)
```

## Generate Realised Webs with Power-Law

Generate 5 hypothetical realised webs

```{r powerlaw-webs}
realised_webs <- powerlaw_prey(
  el = edgelist,
  n_samp = 5,
  y = 2.5
)

# Inspect first realised web
realised_webs[[1]]
```

## Customizing the Power-Law Distribution

```{r custom-func}
# Define a custom in-degree distribution function
custom_func <- function(r, M, y) (M - r + 1)^(-y)

realized_webs_custom <- powerlaw_prey(
  el = edgelist,
  n_samp = 5,
  y = 2,
  func = custom_func
)

# First custom web
realized_webs_custom[[1]]
```

